---
title: "STA457TUT5"
output:
  pdf_document: default
  html_document: default
---

For an AR(2) process, we have that:

$$x_t = \phi_1 x_{t-1} + \phi_2 x_{t-2} + w_t $$ 

In our case, $\phi_1 = 1.5$, $\phi_2 = -0.75$ .

The difference equation for the ACF of the process is:

$$\rho(h)-\phi_1\rho(h-1)-\phi_2\rho(h-2) = 0 $$.

So we have a difference equation:

$$\rho(h)-1.5\rho(h-1)+0.75\rho(h-2) = 0 $$

Solve for general expression of $\rho(h)$. Let $z_1$and $z_2$ be the roots of the associated polynomial:

$$\phi(z) = 1-\phi_1z - \phi_2z^2$$

$$\phi(z) = 1-1.5z + 0.75z^2$$

where $z_1 = 1+i\sqrt3/3$, $z_2 = 1-i\sqrt3/3$.

The general solutions are:

$$\rho(h) = c_1z_1^{-h} + c_2z_2^{-h} $$
For initial conditions, $$\rho(0) = c_1+c_2 = 1$$
$$\rho(-1) = c_1z_1 + c_2z_2 = \phi_1/(1-\phi_2) = 6/7 $$

We solve that :
$c_1 = 1/2+(\sqrt3/14)i$, $c_2 = 1/2-(\sqrt3/14)i$

The general solution is given by:

$$\rho(h) = (1/2+(\sqrt3/14)i)(1+(\sqrt3/3)i)^{-h} + (1/2-(\sqrt3/14)i)(1-(\sqrt3/3)i)^{-h} $$

We use a loop to calculate the given recurrence relation.
  
```{r}
loop <- function(n){
  
  a <- 1
  b <- 6/7
  for (i in 2:n){
    a_new <- b
    b <- 1.5*b - 0.75*a
    a <- a_new
  }
  return(b)
}
```

```{r}
ACF<-c(1,6/7,loop(2),loop(3),loop(4),loop(5),loop(6),loop(7),loop(8),loop(9),loop(10))
X<-c(0:10)
plot(X,ACF,type = "h",col = "blue")
abline(h=0)
```
We use ARMAacf function to check the value we calculated.


```{r}
ACF_r = ARMAacf(ar= c(1.5,-0.75),ma=0,10)
plot(ACF_r,type = "h",xlab = "lag",col = "red")
abline(h=0)
```

