---
title: "STA457TUT8"
author: "Shu Wang"
date: "2022/4/4"
output: pdf_document
---

```{r}

library('astsa')
data(salt)
```

# (a)
```{r}
tsplot(salt,type = 'o',col = 'blue')
```

```{r}
n = length(salt)
par(mfrow = c(2,1),mar = c(3,3,1,1),mgp = c(1.6,0.6,0))
Per = Mod(fft(salt-mean(salt)))^2/n
Freq = (1:n -1)/n
plot(Freq[1:50],Per[1:50],type = 'h',lwd = 3,ylab = 'Periodogram',xlab = 'Frequency')

```
```{r}
y = cbind(1:50,Freq[1:50],Per[1:50])
y<-y[order(y[,3]),]
y
```

We see that 1/freq(5) = 16 day cycle and 1/freq(4) = 21.333 day cycle and 1/freq(8) = 9.143 day cycle are the most prominent periodic components of the data.

# (b)

```{r}
salt.per = mvspec(salt,log = 'no')
abline(v = 0.063,lty = 2)
abline(v = 0.11,lty = 2)
```

We notice a narrow-band peak at $\omega = 0.063\Delta = 5.25* 10^{-3}$,The occurring years is $1/0.063 = 15.87$ years.

```{r}
nextn(length(salt))
```


# (c)



```{r}
P2<-salt.per$details[order(salt.per$details[,3],decreasing = TRUE),]
P2[1,];P2[2,];P2[3,]
```




```{r}
U =qchisq(0.05,2)
L =qchisq(0.95,2)

##90% CIs for the dominant frequencies for saltemp series in part(a)
salt.u1 = 2*P2[1,3]/U
salt.l1 = 2*P2[1,3]/L
salt.u2 = 2*P2[2,3]/U
salt.l2 = 2*P2[2,3]/L
salt.u3 = 2*P2[3,3]/U
salt.l3 = 2*P2[3,3]/L
##Create a data frame for the CIs
Result <- data.frame(Series=c(rep("salt",3)),
Dominant.Freq=c(P2[1,1],P2[2,1],P2[3,1]), Spec=c(P2[1,3],P2[2,3],
P2[3,3]),
Lower=c(salt.l1,salt.l2,salt.l3),
Upper=c(salt.u1,salt.u2,salt.u3))
Result[1:2,3:5] = round(Result[1:2,3:5], 4)
Result

```

We cannot establish the significance of the first peak since the periodogram ordinate is 60.67, which lies in the confidence intervals of the second and third peak.

We cannot establish the significance of the second peak since the periodogram ordinate is 33.49, which lies in the confidence interval of the first and third peak.

We cannot establish the significance of the third peak since the periodogram ordinate is 23.69, which lies in the confidence interval of the first and second  peak.
